### 2022-12-28 paper/paper.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

\usepackage[hyphens]{url}

  \journal{Annals of Emergency Medicine} % Sets Journal name

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
% TODO: Currently lineno needs to be loaded after amsmath because of conflict
% https://github.com/latex-lineno/lineno/issues/5
\usepackage{lineno} % add
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[top=25mm, left=25mm, right=25mm, bottom=25mm,headsep=10mm, footskip=12mm]{geometry}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Probabilistic forecasting of hourly Emergency Department arrivals},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{adjustbox, float,lscape,amsmath,bm}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\begin{frontmatter}

  \title{Probabilistic forecasting of hourly Emergency Department arrivals}
      \cortext[cor1]{Corresponding author}

  \begin{abstract}
  An accurate forecast of Emergency Department (ED) arrivals by an hour of the day is critical to meet patients' demand. It enables planners to match ED staff to the number of arrivals, redeploy staff, and reconfigure units. This can have many advantages for healthcare staff and the quality of care delivered to patients. In this study, we develop an innovative model based on Generalised Additive Models and an advanced dynamic model based on exponential smoothing to generate an hourly probabilistic forecast of ED arrivals for a prediction window of 48 hours. We compare the forecast accuracy of these models against appropriate benchmarks, including TBATS, Poisson Regression, Prophet, and simple empirical distribution. We use Root Mean Squared Error (RMSE) to examine the point forecast accuracy and assess the forecast distribution accuracy using Quantile Bias, PinBall Score and Pinball Skill Score. Our results indicate that the proposed models outperform their benchmarks for point and probabilistic forecasts. Our developed models can also be generalised to forecast hourly arrivals in other services such as hospitals, ambulances, or clinical desk services.
  \end{abstract}
    \begin{keyword}
    Emergency Department, Poisson Regression, Probabilistic Forecasting, Generalised Additive Models, Intermittent Exponential Smoothing \sep 
    Emergency Department, Poisson Regression, Probabilistic Forecasting, Generalised Additive Models, Intermittent Exponential Smoothing
  \end{keyword}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Forecasting Emergency Department (ED) arrivals is critical for informing
staffing and scheduling decisions to meet the needs of patients.
Accurate ED demand forecasts contribute to a better decision-making
process regarding resources allocation and staffing. This is one of the
best ways to optimise resources utilisation and minimise related costs.
An accurate forecast of patient arrivals is crucial in ED services to
depict various courses of action that can result in massive savings in
terms of patient lives. Inability to match the staff with the demand
might result in patients overcrowding the system, which is a severe
problem that causes challenges for the patient flow
\citep{derlet2002overcrowding}. Also, it is related to increasing length of
stay \citep{muhammet2015forecasting}, low patient satisfaction, unexpected
return visits to services, increased health care costs, inaccuracy in
electronic medical records and other \citep{rostami2020anticipating}.

Accurate forecasting of arrivals by the hour of the day enables planners
to match staff to meet anticipated patients, reconfigure units and
redeploy staff. This has many advantages for both patients, staff and
the quality of provided services. Hourly forecasts are required to
inform the short-term operational planning for the current and the
upcoming shifts of the day. This involves the short-term decision-making
related to the execution of the delivery process in ED. The combination
of an hourly arrival forecast, current staff, being occupied, resource
availability and waiting times at ED provide information on the state of
the unscheduled care system across the service. Having this complete
picture enables the delivery managers to focus on the areas that require
intervention to allow the most effective delivery of the service to the
patients. However, compared with lower frequency time series forecasting
such as monthly, quarterly and yearly, hourly forecasts are challenging
because the noise caused by random variation may overshadow any pattern
in the time series. Hourly time series generally exhibit multiple
seasonal cycles of different lengths: hourly, daily, weekly, and yearly.
They may also express nonstationarity, and their profile may change over
time. Therefore, an appropriate forecasting model should consider these
features to accurately predict hourly demand admissions.

There are few studies that look at forecasting hourly arrivals in ED and
other hospital services using historical time series data and/or
predictors such as patient characteristics, weather, holidays and public
events. These studies use multiple approaches, including Exponential
Smoothing \citep{SvetunkovAdam}, Autoregressive Integrated Moving Average
(ARIMA) \citep{hyndman2021forecasting}, Autoregressive Conditional
Heteroskedasticity (ARCH) \citep{bollerslev1994arch}, Vector Autoregressive
model \citep{lutkepohl2013vector}, TBATS \citep{de2011forecasting} and Artificial
Neural Networks \citep{hyndman2021forecasting}. \textcolor{blue}{We have identified a number of limitations and gaps in the area of ED forecasting that motivated us to undertake this study. These gaps and their importance are discussed below.} \textcolor{blue}{Most of these studies are limited to only predicting future arrivals as a point
forecast (a single number), which does not quantify any uncertainty associated with the number of future arrivals.} There are few studies that report uncertainty by presenting prediction intervals, but \textcolor{blue}{there is no study generating and evaluating the entire forecast distribution of arrivals. Reporting the uncertainty via the forecast distribution is potentially valuable in this setting and has practical implications for those managing Emergency Departments because the consequences of inadequate staffing are asymmetric}. This asymmetry arises because it is preferable to incur a small opportunity cost associated with utilised staff rather than compromise service levels if staff levels are insufficient. Probabilistic forecasts inform decision-makers about exposure to these risks and potentially enable those risks to be managed more efficiently. Furthermore, if the impact of under- and overstaffing can be quantified,
probabilistic forecasts allow \emph{optimal} decisions that balance the cost associated with under- and overstaffing to be calculated. Therefore, in this paper, in addition to generating point forecasts, we also produce and evaluate density forecasts of hourly ED arrivals, comparing several methods for this task. \textcolor{blue}{Another drawback of existing studies is that the
datasets used are relatively small (e.g. time period of 1-2 years), making it challenging to report the forecast accuracy using robust approaches such as time series cross-validation. Such results might not be generalisable. Additionally, most of the forecasting methods used in these publications do not consider the full extent of the multiple seasonality of hourly ED arrivals. Moreover, hourly ED time series may contain low volume values and potentially zeros in some hours of the day, which brings some challenges to traditional time series forecasting resulting in producing negative forecasts of future arrivals.
Finally, all previous publications
referenced in this paper are not fully reproducible as underlying data and code are not available.}

In this paper, we aim at filling these gaps and generate forecasts for a prediction window of 48 hours. Our contributions to the literature are summarised as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  We produce probabilistic forecasts, in addition to the point
  estimation, quantifying uncertainties in future hospital admission,
  and comparing different forecasting methods using a suite of
  well-established evaluation metrics;
\item
  We develop an advanced dynamic model to forecast ED arrivals based
  on iETS \citep{Svetunkov2019a} and ETSX models with a modification for
  multiple frequencies, which produced highly-accurate point
  forecasts;
\item
  We develop a novel model to produce a probabilistic forecast of ED
  arrivals based on Generalised Additive Models for Location Scale and
  Shape, which accounts for i) the bounded and non-Gaussian
  distribution of arrivals, ii) multiple seasonalities, weather and
  holiday effects, and iii) variation in forecast uncertainty;
\item
  We benchmark the accuracy of our model against appropriate models
  used when multiple seasonality is present, i.e.~Prophet, TBATS,
  Poisson Regression, Exponential Smoothing State Space model (ETS)
  and the simple empirical distribution of the arrivals;
\item
  We provide data and code enabling reproduction and refinement of the
  proposed approach and benchmarks. The proposed approach could also
  be generalised to forecast hourly requirements for longer horizons
  and in other services \citep{al2021empirical}, such as inpatient and
  outpatient care services, the number of attended incidents in
  ambulance services, or call volumes in clinical desk services.
\end{enumerate}

The rest of the paper is organised as follows: In section \ref{lit}, we
provide a brief overview of hourly ED arrival forecasting; In Section
\ref{data}, we present the hourly time series of an ED arrival and use
various plots to highlight important patterns. In section \ref{model},
we describe the modelling approach and benchmark methods. We then
discuss the performance evaluation metrics in section \ref{accuracy};
in section \ref{result}, we present and discuss our results. Finally,
we summarise our findings and present ideas for future research in
section \ref{conclusion}.

\hypertarget{lit}{%
\section{Research background: hourly ED forecasting}\label{lit}}

\textcolor{blue}{
There is a substantial number of studies that employ models to forecast admissions and arrivals to inform planning and decision making in the healthcare. Areas such as forecasting call volume arrivals, ambulance demand and Emergency Department forecasting has received a significant attention. We refer interested readers to  some extensive reviews of the relevant literature} \citep{mingliterature2022, gul2020review, ibrahim2016modeling}. \textcolor{blue}{The time granularity considered by these studies spans from hourly to yearly across different parts of the healthcare. However, given
the focus of this study, we only discuss hourly ED forecasting.}

\citet{hertzum2017forecasting} used linear regression, ARIMA, and Naïve to
investigate whether accurate hourly accident and emergency department
patient arrivals and occupancy forecasts can be generated using calendar
variables. \citet{hertzum2017forecasting} study found that patient arrival
variation is larger across the hours of the day than across the days of
the week and the months of the year. In terms of the hour of the day,
patient arrivals peaked around noon. For days of the week, Monday is the
busiest day, while weekends are the quietest ones. July-August are the
months with the highest number of patient arrivals, while January and
February have the lowest numbers. They indicate that regression and
ARIMA models performed similarly in modelling patient arrivals, while
ARIMA outperformed regression models in modelling accident and emergency
department occupancy.

\citet{choudhury2020forecasting} used ARIMA, Holt-Winters, TBATS, and neural
network methods to forecast hourly accident and emergency department
arrivals. ARIMA model was selected as the best fit model. Authors
claimed that ARIMA provided high and acceptable hourly ED forecasting
accuracy, even outperforming TBATS. \citet{Cheng2021} developed an ARIMA model
for ED occupancy with a seasonal component and exogenous variables,
which outperformed a rolling-average benchmark. They also produce
prediction intervals, a form of the probabilistic forecast, which were
found to be well-calibrated, a necessary property for such forecasts.

\citet{morzuch2006forecasting} used the Unobserved Components Model (UCM), in
which each component of the time series is separately modelled as
stochastic. Double-seasonal exponential smoothing and standard
Holt-Winters were used to forecast ED arrival for a horizon of 168
hours. The hourly data collected from an ED in Pennsylvania showed no
trend and two seasonal cycles: a within-day and a within-week seasonal
cycles. The double seasonal model recorded lower RMSEs for all the 168-hour horizons, which was expected due to the strong hourly
seasonality of the time series.

\citet{mccarthy2008challenge} employed a Poisson log-linear regression model,
including independent variables such as temporal factors (e.g.,
hour-of-day, day-of-week, type-of-day, season, and calendar year),
patient characteristics (i.e., age, gender, insurance status, triage
level, mode of arrival, and ambulance diversion status) and climatic
factors (i.e., temperature and precipitation) to model patient demand
for ED services. The authors produced probabilistic predictions in the
form of \(50\%\) and \(90\%\) prediction intervals for the number of hourly
arrivals. Hourly data of ED arrivals in the 1-year study period was
modelled and analysed, and it was suggested that the model could be used
for forecasting. However, model evaluation was performed in-sample on
only one year of data, so it is unclear how this approach would perform
in a forecasting setting or compare to simpler approaches. However, the
length of the time series in this study was very short (only one year),
which did not allow for a rigorous out-of-sample evaluation.

\citet{schweigler2009forecasting} investigated whether time series methods
could accurately generate short-term forecasts of ED bed occupancy. A
year-long dataset of hourly ED bed occupancy was collected from three
facilities. The authors implemented an hourly historical average model,
SARIMA model, and sinusoidal model with autocorrelated error for each
facility. The historical average model was based on the mean occupancy
for each site, for each hour of the day, while the sinusoidal model was
based on four parameters: an AR term, a sine coefficient, a cosine
coefficient and an intercept. They evaluated the forecast accuracy of
four and twelve hours forecast horizons using RMSE. They found that both
SARIMA and the sinusoidal models outperformed the historical average
(for example, at site 2, the two models improved by \(33\%\) the 12-hour
forecasts generated by the historical average).

\citet{kim2014predicting} compared different univariate and multivariate time
series forecasting techniques to predict patient volume for a Hospital
Medicine programme. The study evaluated linear regression, exponential
smoothing, ARIMA, SARIMA, Generalized Autoregressive Conditional
Heteroskedasticity (GARCH) and Vector Autoregressive (VAR) models to
forecast for 4 and 24 hours ahead. They used Mean Absolute Percentage
Error (MAPE) to report the forecast accuracy. The authors found that the
ARIMA outperformed all the other models.

Table \ref{tab:summarylit} summarises the relevant papers.

\begin{landscape}\begin{table}[!h]

\caption{\label{tab:summarylit}Summary of studies in forecasting hourly arrivals in Emergency Department}
\centering
\resizebox{\linewidth}{!}{
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{lr>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{4em}l>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{8em}ll}
\toprule
Reference & Year & Variable & Horizon & Length of dataset & Method & Metric & Probabilistic & Seasonality\\
\midrule
Current study & 2023 & ED arrivals & 48h & 5 years & Naïve; Poisson \& Quantile Regression; Exponential Smoothing; Prophet; TBATS;    Gradient Boosting ; ADAM; GAM & RMSE, Pinball score, skill score, Quantile bias & YES & YES\\
Cheng et al. & 2021 & ED visits & 1h to 4h & 1 year & SARIMAX; Holt-Winters; VAR; ARIMA & MSE, MAE, MAPE, Prediction interval coverage & NO & Single\\
Choudhury and Urena & 2020 & ED arrivals & 1h to 24h & 4 years & ARIMA; Holt-winters; TBATS; ANN & RMSE, ME & NO & Multiple\\
Asheim et al. & 2019 & ED arrivals & 3h & 5 years & Poisson regression & MAPE & NO & Single\\
Hertzum & 2017 & ED arrivals & 1,2,4,8,24h & 3 years & linear regression; SARIMA; Naïve & MAE, MAPE, 
MASE & NO & Single\\
Kim et al. & 2014 & Hospital admission & 4h, 24h & 3 years & Linear regression; Exponential smoothing; ARIMA; GARCH; VAR & MAPE & NO & Single\\
Cote et al. & 2013 & ED arrivals & 24h & 2 years & Fourier regression & R\textasciicircum{}2,  Standard Error & NO & Single\\
Chase et al. & 2012 & ED CUR & 30m
1h, 2h, 4h, 8h, 12h & 1 year & Binary regression & NA & NO & Single\\
Schweigler et al. & 2009 & Bed occupancy & 4h, 12h & 4 years & Hourly historical average; SARIMA; Sinusoidal model with autocorrelated error & RMSE & NO & Single\\
Jones et al. & 2009 & ED census & 24h & 2 years & VAR;  Holt winters & MAE & NO & Single\\
McCarthy & 2008 & ED arrivals & n/a & 1 year & Poisson log-linear regression model & Prediction interval coverage & Partially & Multiple\\
Channouf et al. & 2007 & Ambulance admission & 1h, 3h, 6h, 12h, 13h, 14h, 17h, 23h, 24h & 2 years & Regression & RMSE & No & Single\\
Morzuch and Allen & 2006 & ED arrivals & 168h & 3 years & Double Exponential Smoothing; Additive Holt Winter & RMSE & No & Multiple\\
\bottomrule
\end{tabular}}
\end{table}
\end{landscape}

\citet{asheim2019real} developed a Poisson time-series regression model with
continuous day-of-week and week-of-year effects to implement a real-time
system that could forecast ED arrivals on 1, 2, 3 hours ahead. Measuring
the accuracy using the MAPE, \citet{asheim2019real} noticed that significant
improvement happened when the time of notification was incorporated into
the model, especially in the one-hour horizon.

\citet{cheng2021forecasting} used one year of ED visits time series to evaluate
the Rolling Average, SARIMAX, ARIMA, VAR and Holt-Winter to forecast ED
occupancy up to 4-hours ahead. The forecast accuracy is evaluated using
Mean Squared Error (MSE), Mean Absolute Error (MAE) and MAPE for point
forecast and coverage for prediction intervals of \(80\%\) and \(95\%\).
They show that SARIMAX provides a more accurate forecast of hourly ED
occupancy.

According to the studies mentioned above, it can be said that they have
shown complications in forecasting hourly patient accident and emergency
department visits, and the application of forecasting hourly patient
visits is not well established. Some of the studies claimed that the
accuracy of forecasting models on hourly accident and emergency
department data is low \citep{boyle2012predicting, hertzum2017forecasting},
while others mentioned that the accuracy of ED hourly forecast is at an
acceptable level \citep{choudhury2020forecasting, mccarthy2008challenge, schweigler2009forecasting}.

There are a few limitations in the literature which encourage us to
undertake this research and examine different forecasting approaches:

\begin{enumerate}
\def\labelenumi{(\roman{enumi})}
\tightlist
\item
  Current approaches to forecast hourly ED arrivals do not fully
  consider the feature of data such as multiple seasonal cycles and
  changing profile over time;
\item
  Almost all research studies produce point forecasts and, at best,
  report prediction intervals. There is a lack of studies presenting
  the entire forecast distribution of hourly ED arrivals that better
  represent the uncertainty of future arrivals, providing a holistic
  picture of future demand for a planner;
\item
  most studies are not reproducible, as it is almost impossible to
  reapply the approaches without the help of the authors of those
  papers;
\item
  studies are limited in terms of the length of historical data used
  for training purposes and forecast performance evaluation and
\item
  some studies in this area lack a rigorous experimental design, i.e.
  they do not use benchmark methods or report forecast accuracy.
\end{enumerate}

\hypertarget{data}{%
\section{Preliminary analysis}\label{data}}

Data used in this study comprises counts of patients' arrival times at
one of the largest ED units in the UK between April 2014 and February
2019, extracted from the ED administrative database of the hospital. We
aggregated the patients' arrival times to obtain hourly arrivals, which
are used in this study. Figure \ref{fig:hourly-plot-ridge} illustrates
the distribution of arrivals for each hour of the day and the day of the
week. Although the data is noisy, it reveals some systematic patterns.

\begin{figure}[H]

{\centering \includegraphics{paper_files/figure-latex/hourly-plot-ridge-1}

}

\caption{Distrubution of admission by hour-of-day and day-of-week. Most days have a distinct pattern, such as relatively high arrivals on Monday mornings and in the early hours of Saturday and Sunday morning.}\label{fig:hourly-plot-ridge}
\end{figure}

It is clear that the number of arrivals has a sub-daily structure,
similar to the one summarised by \citet{hertzum2017forecasting}. The ED
arrivals decrease between midnight and early morning and then increase
until the evening, decreasing after that again. It is also clear that ED
service gets systematically more visits on Mondays between 8 a.m. and 5
p.m. Moreover, the number of arrivals around midnight is slightly higher
for Saturdays and Sundays.

Figure \ref{fig:hourly-plot-ridge} also highlights significant skewness
for almost every hour of the day that varies with time-of-day, which
should be accounted for in forecasting methods. Some skewness might be
related to holidays and special events. It is also clear that arrivals
are less volatile between midnight and early morning.

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/seasonplot-dofw-1}

}

\caption{Subseries plot: day of weekly arrivals}\label{fig:seasonplot-dofw}
\end{figure}

Figure \ref{fig:seasonplot-dofw} \textcolor{blue}{illustrates the daily subseries plot, with the x-axis representing the date and y-axis the ED arrivals. Each individual plot illustrates how arrivals change over time for a each day of week (i.e. seven days) from Monday to Saturday. The blue line shows the average arrival for the given day. It is clear that ED arrivals on Monday are higher followed by Saturday. This indicates that there are
significantly more arrivals on Mondays and Saturdays compared to the rest of the week. This might be due to the closure of General Practitioners outpatient clinics over the weekend.}

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/seasonplot-weekofyear-1}

}

\caption{Arrivals by week-of-year. There is an annual trend of reduced arrivals durnng the summer and winter holiday period.}\label{fig:seasonplot-weekofyear}
\end{figure}

Figure \ref{fig:seasonplot-weekofyear} highlights the week of year
seasonality in the ED arrivals. We observe that arrivals are
significantly lower from week 29 to 35, corresponding to the Summer
period. Moreover, the number of arrivals is lower at the beginning and
the end of the year. The arrivals increase from week 36 and peak in
weeks 39-42. This corresponds to the September - October period.

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/date-plot-1}

}

\caption{Daily arrivals over the entire dataset with moving average trend (blue line) and dates of outliers labled. Low outliers invariables correspond to public holidays.}\label{fig:date-plot}
\end{figure}

Finally, Figure \ref{fig:date-plot} presents the time plot of daily
arrivals. Each point represents one day, and points are shape-coded by
day-of-week to show the weekly cycle. The figure shows more arrivals on
Mondays than on other weekdays, agreeing with the previous findings.
Moreover, we can observe a long-term trend line and the significant
effect of some holidays. We can see that arrivals near Christmas and New
Year's day are significantly lower than other days of the year.
Moreover, Figure \ref{fig:date-plot} allows us to identify the impact
of special events on arrivals. For instance, we see that arrivals are
significantly low for 01-03 of March 2018. These days correspond to the
Storm Emma with heavy snowfall that resulted in travel disruption, mass
power outages, and schools closed in the UK \citep{stormemma2018}.

Based on this analysis and the literature review, we should consider
models that can take the following into account:

\begin{itemize}
\tightlist
\item
  Hour-of-day, day-of-week, and week-of-year seasonalities,
\item
  Long-term trend (or a slowly changing level),
\item
  Calendar events, such as holidays,
\item
  Lags of calendar events to accommodate the potential changes in
  demand the next day after a holiday,
\item
  Other events, such as sporting fixtures,
\item
  Temperature effects.
\end{itemize}

We propose several forecasting models that account for the structures
outlined above.

\hypertarget{model}{%
\section{Model building}\label{model}}

\hypertarget{naive}{%
\subsection{Naive}\label{naive}}

We start with one of the simplest forecasting approaches used in
practice - assuming that in the next few hours, everything will be the
same as in the similar hours of a similar day in the past. This is
called ``Naïve''. In our case, given that we need a distribution of
values, we will use a modified approach, where the
empirical distribution of the hourly arrival time series is used to
forecast the future arrival distribution \citep{la2021new}. We consider the
empirical distribution of all available historic data (Benchmark-1) and
the empirical distribution of the most recent year of historic data on a
rolling basis (Benchmark-2) to capture potential changes in behaviour
over time.

\hypertarget{poisson-regression}{%
\subsection{Poisson Regression}\label{poisson-regression}}

Regression is one of the most popular forecasting methods that use
explanatory variables to predict a variable of interest (in our case,
the ED arrivals). The classical linear regression model is formulated as
\begin{equation}
  {y}_t = \mathbf{x}_t' \boldsymbol{\beta} + \epsilon_t ,
\label{eq:linearRegression}
\end{equation} where \(\mathbf{x}_t\) is the vector of explanatory
variables, \(\boldsymbol{\beta}\) is the vector of parameters,
\(\epsilon_t\) is the error term, which is typically assumed to follow
Normal distribution with zero mean and a fixed variance, and \(t\) is the
time index. However, in the context of healthcare and ED arrivals, the
assumption of Normality is unrealistic because the number of admitted
patients is an integer and non-negative. So the linear regression model
should be substituted by some other model. One of the models that is
frequently used in practice is the Poisson regression \citep[see, for example,][]{mccarthy2008challenge}, which can be summarised as \begin{equation}
  \begin{aligned}
    & {y}_t \sim \mathrm{Poisson} \left( \lambda_t \right) \\
    & \log \lambda_t = \mathbf{x}_t' \boldsymbol{\beta}
  \end{aligned} \quad .
\label{eq:PoissonRegression}
\end{equation}

The logarithm in \eqref{eq:PoissonRegression} is needed to ensure that
the parameter of Poisson distribution is always positive. This model can
be estimated via maximisation of the likelihood function based on
Poisson mass function . There is no single
correct answer when selecting explanatory variables for the model, and
the decision needs to be made for each specific case. In our experiment,
we will only include dummy variables, capturing a variety of calendar
events:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Hour of the day,
\item
  Day of the week,
\item
  Week of the year,
\item
  Holidays (such as Christmas, New Year etc.),
\item
  24 hours lags of holidays.
\end{enumerate}

The variables \emph{1--3} allow modelling the seasonal patterns on the
appropriate level of detail throughout the year, while \emph{4} covers the
changes in admittance due to calendar events. Finally, \emph{5} is needed to
capture the potential phenomenon of change in admittance after the
holiday (e.g.~people might try not to go to the hospital on Christmas
eve and thus go the next day). This model assumes that all these effects
are deterministic and do not change over time. Still, the exponentiation
in \eqref{eq:PoissonRegression} introduces an interaction effect between
dummy variables so that the 3 pm on Monday in January will be different
from 3 pm on Monday in July, although the parameters for the hour of day
and day of the week are fixed and do not change over time. We use the
\texttt{alm()} function from the \texttt{greybox} package \citep{Svetunkov2021Greybox} for
R \citep{RTeam2021} for the experiments and denote this model as ``Poisson
Regression''.

\hypertarget{ets---exponential-smoothing}{%
\subsection{ETS - Exponential Smoothing}\label{ets---exponential-smoothing}}

\citet{Hyndman2008b} developed a state space approach for exponential smoothing
models. The model can have a set of elements, including different types
of Error, Trend and Seasonal components (thus \emph{ETS}). Given the
popularity of the ETS model in the forecasting community, we decided to
include the basic ETS(A,N,A) model with the seasonal component with a
frequency of 24 (hour of the day) as a benchmark. This was done using
the \texttt{adam()} function from the \texttt{smooth} package \citep{Svetunkov2021Smooth}
for R and denoted as \emph{ETS}. This model does not capture the day of week
or week of year effects, does not include explanatory variables, but its
seasonal component and level change over time. This model is included as
a benchmark, only to see how the other models perform compared to it.

\hypertarget{prophet}{%
\subsection{Prophet}\label{prophet}}

Prophet is a forecasting procedure created by Facebook
\citep{taylor2018forecasting} that accounts for multiple seasonality,
piecewise trend and holiday effects. Prophet is robust to missing data
and shifts in the trend and typically handles outliers well. Prophet
works well on daily data seen in Facebook. It is robust and automated,
making it easy to learn for beginners. The implementation may be less
flexible than other methods. The model itself relies on the Multiple
Source of Error state space model, initially proposed by \citet{Kalman1960}.
The model is incorporated using the corresponding implementation of the
Fable package in R. We use the \texttt{prophet()} function from the \texttt{fable}
package \citep{fable2020}. Note that the input data is assigned with an
hourly and daily seasonality. \textcolor{blue}{This method has been adopted in some healthcare service providers in the United Kingdom to produce forecasts, therefore we have included it as one of our benchmarks.}

\hypertarget{tbats}{%
\subsection{TBATS}\label{tbats}}

\citet{de2011forecasting} proposed a model to deal with time series exhibiting
multiple complex seasonalities called ``TBATS''. It includes a Box-Cox
Transformation, ARMA model for residuals and a trigonometric expression
of seasonal terms. The latter gives the model more flexibility to deal
with fractional seasonality and reduces the parameters of the model when
the frequencies of seasonalities are high. We fit a TBATS model using
the \texttt{tbats()} function from the \texttt{forecast} package in R
\citep{forecastpackage2020}.

\hypertarget{quantile-regression-and-gradient-boosting-machines}{%
\subsection{Quantile Regression and Gradient Boosting Machines}\label{quantile-regression-and-gradient-boosting-machines}}

Quantile regression allows the production of density forecasts without
assuming a fixed distributions shape controlled by a small number of
parameters, such as the Poisson distribution with parameter \(\lambda\).
By producing forecasts of multiple quantiles the full predictive
distribution can be constructed. Quantile regression is particularly
useful where data do not follow a simple distribution, the
distribution shape changes over time with some covariate, which may
be the case with ED arrivals.

Gradient Boosting Machines (GBM) are a tree-based machine learning model
for regression and classification. Here we produce probabilistic
forecasts of ED arrivals via multiple quantile regression using GBMs as
implemented in the R package \texttt{gbm} \citep{Greenwell2020}. GBMs are a
best-in-class algorithm for similar regression problems characterised by
modest volumes of training data and possible interactions between input
features, and are therefore an appealing choice for ED arrival
forecasting. Here GBMs are fit with the following features: hour of the
day, day of the week, school holiday, temperature and day of the year.
Hyperparameters are chosen via grid search on the training data, and
were chosen as follows: 500 trees, an interaction depth of 2, and a
shrinkage 0.1. An advantage of GBMs is their ability to learn
interactions in comparison to the additive models presented later which
require the user to specify possible interactions between inputs. Here
GBM is included as a reference for the performance of `out-of-the-box'
machine learning models applied with minimal effort. We also explored
the possibility of performing linear regression with additive models,
but these did not perform as well as GBM so are omited for brevity.

\hypertarget{adam-multiple-seasonal-ietsx}{%
\subsection{ADAM: multiple seasonal iETSX}\label{adam-multiple-seasonal-ietsx}}

\citet{SvetunkovAdam2021} proposed a framework for dynamic models called the
Augmented Dynamic Adaptive Model (ADAM). This framework encompasses
ARIMA \citep{Box1976}, ETS \citep{Hyndman2008b} and regression, supporting
multiple frequencies, non-normal distributions and intermittent demand
\citep{Svetunkov2019a}. Based on this framework, we use the ETS(M,N,M) model
with frequencies 24 (hour of the day) and 168 (hour of the week), adding
dummy variables for the week of the year, holidays and lagged holidays.
This way, we update the hour of day and day of week seasonal indices,
keeping the week of year one fixed, thus reducing the number of
estimated parameters. Given that the data exhibits randomly occurring
zeroes, we use the direct probability model of \citet{Svetunkov2019a} to treat
those values. Finally, given the skewness of the empirical distribution
observed in the preliminary analysis, we use the Gamma distribution for
the error term. This model can be formulated as a set of the following
equations: \begin{equation}
    \begin{aligned}
      & y_t = o_t z_t \\
        & \log z_t = \log l_{t-1} + \log s_{1,t-24} + \log s_{2,t-168} + \mathbf{x}_t' \boldsymbol{\beta} + \log \left(1 + \epsilon_{t} \right) \\
        & \log l_{t} = \log l_{t-1} + \log( 1  + \alpha \epsilon_{t}) \\ 
        & \log s_{1,t} = \log s_{1,t-m} + \log( 1  + \gamma_1 \epsilon_{t}) \\
        & \log s_{2,t} = \log s_{2,t-m} + \log( 1  + \gamma_2 \epsilon_{t}) \\
        & o_t \sim \text{Bernoulli} \left(\mu_{a,t} \right) \\
        & a_t = l_{a,t-1} \left(1 + \epsilon_{a,t} \right) \\
        & l_{a,t} = l_{a,t-1}( 1  + \alpha_{a} \epsilon_{a,t}) \\
        & \mu_{a,t} = \min(l_{a,t-1}, 1)
    \end{aligned} ,
    \label{eq:ADAMModel}
\end{equation} where \(\alpha\), \(\beta\), \(\gamma_1\), \(\gamma_2\) and
\(\alpha_a\) are the smoothing parameters, defining how adaptive the
components of the model should be, \(l_t\) is the level component for the
demand sizes, \(s_{1,t}\) and \(s_{2,t}\) are the seasonal components,
\(\boldsymbol{\beta}\) is the vector of parameters for the explanatory
variables, \(o_t\) is the binary variable, which is equal to one, when
demand occurs and to zero otherwise, \(l_{a,t-1}\) is the level component
for the occurrence part of the model, and
\(\left(1+\epsilon_t \right) \sim \Gamma(s^{-1}, s)\), where
\(s=\frac{1}{T} \sum_{t=1}^{T} e_{t}^2\) is the scale of the distribution.
Finally, \(a_t\) is an unobservable series, underlying the occurrence part
of the model and \((1 + \epsilon_{a,t})\) is an unobservable error term
for \(a_t\). \citet{Svetunkov2019a} discuss how to estimate such a model. We
expect this model to perform on par with the Poisson regression,
potentially outperforming it in some instances, due to the dynamic
nature of the model (level and seasonal components). Although the data
is integer-valued, we expect that Gamma distribution will be a good
approximation for it. If integer-valued quantiles are needed, then
rounding up can be done for them (see Appendix \ref{quantilesceiling}
for the explanation). This model is implemented in the \texttt{adam()} function
from the \texttt{smooth} package \citep{Svetunkov2021Smooth} for R and is denoted in
our experiment as ``ADAM-iETSX''.

\hypertarget{gamlss}{%
\subsection{GAMLSS}\label{gamlss}}

Suppose we assume that our predictive distribution follows a given
parametric distribution, as in Poisson regression discussed above. In
that case, the forecasting task becomes one of predicting the future
values of that distribution's parameters. We then can use Generalised
Additive Models for Location, Scale and Shape (GAMLSS). These are the
distributional regression models where the parameters are modelled as
additive function of explanatory variables. This provides a powerful
and flexible framework for probabilistic forecasting, provided that
suitable distribution and additive model structures can be found. In
practice, this means employing expert judgement and experimenting with
various distributions and evaluating their suitability using available
training data.

Let \(F_t(y_t)\) be a predictive cumulative probability distribution of
\(y_t\). In a distributional regression context, \(F_t(y_t)\) is modelled
via a parametric model, \(F(y_t|\bm \theta_t)\), where \(\bm \theta_t\) is
an \(m\)-dimensional vector of parameters. In a GAMLSS framework of
\citet{Rigby2005} the elements \(j=1,...,m\) of \(\bm \theta_t\) are modelled as
\begin{equation}
    g_j(\theta_{j,t})=\mathbf{A}_{j,t} \bm{\beta}_j + \sum_{i} f_{j,i}({\bm x}^{S_{j,i}}_t), \;\;\; \text{for} \;\;\; j = 1, \dots, m,
    \label{eq:basicGAM}
\end{equation} where \(g_j\) is a monotonic link function,
\(\mathbf{A}_{j,t}\) is the \(t\)-th row of the design matrix
\(\mathbf{A}_j\), \(\bm \beta_j\) is a vector of regression coefficients,
\(\bm x_t\) is a \(d\)-dimensional vector of covariates and
\(S_{j,i} \subset \{1, \dots, d\}\) \textbf{is \ldots{}}. If \(S_{j,i} = \{1, 3\}\),
then following our notation \({\bm x}_{t}^{S_{j,i}}\) is a two dimensional
vector formed by the first and third elements of \(\bm x_t\). Each
\(f_{j,i}\) is a smooth function, constructed as \begin{equation}
    f_{j,i}(\bm x^{S_{j,i}}) = \sum_{k=1}^{K_{j,i}} b^{ji}_k (\bm x^{S_{j,i}}) \beta_k^{ji},
    \label{eq:smmothfunction}
\end{equation} where \(b^{ji}_k\) are spline basis functions of dimension
\(\vert S_{j,i} \vert\), while \(\beta_k^{ji}\) are regression coefficients.
The smoothness of each \(f_{j,i}\) is controlled via ridge penalties, the
definition of smoothness being dependent on the type of effect and
penalty being used. See \citet{Wood2017} for a detailed introduction to
\emph{GAM/GAMLSS} models, smoothing splines bases and penalties.

As our data are counts, the natural starting point is the Poisson
distribution with an additive model for \(\log \lambda_t\) of the form
\begin{equation}
  \log(\lambda_t) = \sum_{i=1}^7 \beta_i \delta(D_i(t)-i) + \sum_{j=1}^7 D_j(t) f_j(H(t)) + t f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad .
 \label{eq:additivemodel}
\end{equation}

The functions \(H(t)\), \(D(t)\) and \(Y(t)\) return the hour of the day
(1--24), day of the week (1--7), and day of the year (1-366) at time
\(t\), respectively, and \(C_t\) is the temperature at time \(t\). This model
is called Poisson-1 in discussions below.

However, experiments on the training data reveal that calibration of
forecasts based on the Poisson distribution is poor, suggesting that the
shape of the distribution is unsuitable for the present application. In
particular, we observe that forecast uncertainty appears to vary
depending on the time of day and possibly other explanatory variables.
Therefore, we consider more flexible, two-parameter distributions to
specify additive models for both location and scale parameters,
specifically the truncated Normal distribution, with truncation at 0.
The resulting density forecasts are given by \begin{equation}
F_t(y_t,\mu_t,\sigma_t) =  \frac{\Phi\left( \frac{y_t-\mu_t}{\sigma_t} \right) - \Phi\left( \frac{-y_t}{\sigma_t} \right)}{1 - \Phi\left( \frac{-y_t}{\sigma_t} \right)}
\label{eq:truncatedn}
\end{equation} with additive models
\begin{align*}
    & \mu_t = \sum_{i=1}^{10} \beta_i D^{+}_i(t) + \sum_{j=1}^{10} D^{+}_j(t) f_j(H(t)) + t f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \\
    & log(\sigma_t) = \sum_{i=1}^{10} D^{+}_i(t) f(H(t))
\end{align*} for the mean and variance parameters. This model is
referred to as NOtr-1 below.

Furthermore, we consider an extension to the additive models for
\(\lambda_t\) and \(\mu_t\) above to incorporate school and public holidays
into \(D\). These models are labelled Poisson-2 and NOtr-2. We also
performed experiments with the truncated \(t\) distribution (Ttr-2) and
negative binomial distribution (NBI-2), but these did not result in
forecasts as well calibrated as the truncated normal.

\hypertarget{accuracy}{%
\section{Forecast performance evaluation}\label{accuracy}}

In order to assess the performance of models, we evaluate predictive
quantiles at probability levels 0.05 to 0.95 in steps of 0.05 and
conditional expectations for 0 to 48 hours ahead produced by each model.
We forecast up to 48 hours because this is the operational horizon in
the ED, for which it is possible to make short-term changes in the
shifts for nurses and doctors. The forecasts are produced every 12 hours
for the holdout of 365 days in a rolling origin fashion \citep{Tashman2000},
resulting in 727 origins. Based on these values, several error measures
are calculated to evaluate the performance of models in terms of
specific quantiles and expectation. The latter is measured via Root Mean
Squared Error (RMSE): \begin{equation}
  \mathrm{RMSE} = \sqrt{\frac{1}{h} \sum_{j=1}^h e_{t+j}^2} ,
  \label{eq:RMSE}
\end{equation} where \(h\) is the forecast horizon and \(e_{t+j}\) is the
point forecast error \(j\) steps ahead.

The objective of density forecasts is to be as sharp as possible while
remaining reliable/calibrated \citep{gneiting2007probabilistic}. A forecast
is said to be sharp if the predictive distribution has a relatively
small spread, indicating low uncertainty, which is valuable to
decision-makers provided the forecast is calibrated. Calibration, also
called reliability, is the property that forecast probabilities match
the observed frequency of realisations. If a forecast is calibrated,
then, for example, \(20\%\) of observations should fall below the
\(\alpha=0.2\) predictive quantile (with some tolerance based on the
finite sample size). This property is necessary for forecast
probabilities to be used in quantitative decision-making. Calibration is
typically evaluated visually using reliability diagrams, which plot the
nominal coverage, \(\alpha\), against observed frequency mean
(\(\mathbf{1}(y_{t}\leq q_{\alpha,t})\)). We use several scores to assess
the quantile performance of models.

First, to measure quantile performance, we need to calculate the pinball
score, which is a strictly proper score used to evaluate quantile
forecasts and is the discrete form of the Continuous Rank Probability
Score \citep{hyndman2021forecasting}. It rewards sharpness and penalises
miscalibration, so it measures overall performance. However, calibration
should still be verified separately. Furthermore, The Pinball Score for
an individual quantile matches the loss function minimised in a quantile
regression model. The Pinball Score is given by \begin{equation}
    \text{Pinball} = 
    \frac{1}{T|\mathcal{A}|} \sum_{\alpha \in \mathcal{A}} \sum_{t=1}^T
 \left(q_{\alpha,t} - y_{t} \right)
 \left(\mathbf{1}(y_{t}\leq q_{\alpha,t})-\alpha \right) ,
 \label{eq:pinball}
\end{equation} where \(\mathcal{A} = \{0.05,0.1,...,0.95\}\) is the set of
quantiles being estimated.

To compare model performance, and the significance of any apparent
difference in performance, we will use skill scores, which can be
calculated for any metric via: \begin{equation}
  \mathrm{Skill} = \frac{M_\mathrm{ref} - M}{M_\mathrm{ref}} \label{eq:skillscore}
\end{equation} where \(M\) is the metric's value for the method being
considered, \(M_\mathrm{ref}\) is the metric's value for a reference
method. The skill score shows by how many per cent the reference
approach is worse than the one under consideration. We will use
bootstrap re-sampling of skill scores to determine if the differences in
forecast performance (i.e.~positive or negative skill) are significantly
different from zero \citep{Efron1981Bootstrap}. Here we use the best
performing simple benchmark, Naive (explained in Subsection
\ref{naive}), as the reference model and employ a block-bootstrap
with blocks of length 24h to account for temporal correlations of the
underlying data \citep{hongyi1996bootstrapping, Bergmeir2016303}.

Finally, we have calculated the computational time for one iteration on
the first rolling origin to compare the speed of each function. All
functions were re-estimated on each iteration. ADAM and Poisson
regression estimated the parameters, taking the ones obtained in the
initial model application to the data in the first origin as the
pre-initials. This allowed to speed up the computation for these two
models. The initial estimation of ADAM took approximately one hour and
25 minutes. Each step in the experiment took the time shown in Section
\ref{result}.

\hypertarget{result}{%
\section{Results}\label{result}}

The data is portioned into training (from 2014-04-01 to 2018-02-28) and
test (from 2018-03-01 to 2019-02-28) sets, with all model development
and hyper-parameter tuning performed using training data only. The
rolling origin advances in 12-hour steps and the forecast horizon is set
for 48 steps ahead.

Figure \ref{fig:Pinball} presents Pinball score aggregated across
forecasting horizons for each quantile. It shows that the difference in
performance among the models mainly comes from the middle of the
distribution and somewhat from the upper tail. There is very little
difference in performance for the lower tail (except for TBATS, which
has a consistently higher Pinball value than the other models). This is
interesting and reassuring that the better models are better at
probabilities that matter more to decision-makers.

Probabilistic forecasts are evaluated following the principle of
\emph{sharpness subject to calibration}, meaning that the sharper forecast is
prefered provided that it is calibrated. Mis-calibrated forecasts are
unsuitable for use in decision-making, so they should be excluded.
Calibration is evaluated visually in Figure \ref{fig:quantile-bias},
which highlights a systematic negative bias across all probability
levels in many models, with only the truncated normal and \(t\) family
GAMLSS models (NOtr-1, NOtr-2, Ttr-2) and ADAM-iETSX models showing good
calibration across most probability levels. \textcolor{blue}{ Notably, both benchmarks
exhibit negative quantile bias as they struggle to capture the long term trend of increasing arrivals. This could result in poor staffing decisions. This is because the empirical distribution of whole data fails to characterise how arrivals in ED may change over time.}

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/Pinball-1}

}

\caption{Pinball loss values by quantile. Benchmark-2, NBI-2 and NOtr-2 have similar performance with most improvement of the latter two over the benchmark comming from the lower quantiles. All other models have a greater Pinball score (worse performance) than Benchmark-2 across all quantiles.}\label{fig:Pinball}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/quantile-bias-1}

}

\caption{Quantile bias illustrates the difference between the nominal exceedance of prediceted quantiles and the observed frequency of exceedance. If a predictive model is calibrated, quantile bias should be approximately zero accross all probability levels. Here we see that the predictive distributions produced by TBATS are underdispersed (too narrow/over confident); those produced by NBI-2, GBM-2 and Benchmark-2 bias and underpredict accross all probability levels; and that forecasts produced by ADAM-iETSX, NOtr-2 and Prophet are well calibrated.}\label{fig:quantile-bias}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/lead-time-rmse-1}

}

\caption{RMSE values over the forecast horizon for forecasts issues at midnight. Attendance during the morning is more challenging to predice resulting in greater RMSE during these hours for all models.}\label{fig:lead-time-rmse}
\end{figure}

Figure \ref{fig:lead-time-rmse} reports the RMSE for each forecast
horizon. It illustrates the times of day that are harder to predict --
morning pick-up and afternoon peak. We can see that some models perform
much better than others on specific forecast horizons. For example,
Benchmark does an excellent job in predicting 1 - 5 steps ahead ED
arrivals and, in general, doing well in forecasting arrivals in the
night. In fact, the Benchmark model performs consistently well in terms
of the conditional mean, not making as huge mistakes as, for example,
Prophet, TBATS and GBM do for some lead times.

\begin{table}[!h]

\caption{\label{tab:tab-results}Summary of studies in hourly emergency care forecasting}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{lrrrr}
\toprule
Method & Quantile Bias & Pinball & RMSE & Time (minutes)\\
\midrule
NOtr-1 & 0.0098967 & 1.222583 & 0.2675957 & 451.6620471\\
ADAM-iETSX & 0.0104673 & 1.417260 & 0.0896228 & 92.9348605\\
NOtr-2 & 0.0118522 & 1.208561 & 0.2675957 & 86.5895462\\
Ttr-2 & 0.0140221 & 1.210108 & 0.3324146 & 956.5532849\\
Prophet & 0.0193799 & 1.447037 & 0.2955460 & 20.6755021\\
ETS & 0.0194389 & 1.434862 & 0.0121247 & 10.7175205\\
Poisson-1 & 0.0372137 & 1.204920 & 0.0095263 & 1.4763353\\
Poisson-2 & 0.0373884 & 1.188109 & 0.0082932 & 5.0768588\\
NBI-2 & 0.0540725 & 1.206241 & 0.3830272 & 1.2791870\\
Benchmark-2 & 0.0557392 & 1.217429 & 0.2592800 & 0.0947247\\
GBM-2 & 0.0600153 & 1.261690 & 1.7770897 & 602.4317496\\
TBATS & 0.0855702 & 1.536080 & 0.4859770 & 273.0558176\\
Regression-Poisson & 0.0929416 & 1.293524 & 0.8490258 & 67.1401641\\
Benchmark-1 & 0.1047874 & 1.254491 & 1.0042634 & 0.3874450\\
\bottomrule
\end{tabular}
\end{table}

In above figures (Figures \ref{fig:Pinball}-
\ref{fig:lead-time-rmse}), we only present the performance of the top
seven methods. However, we have evaluated the performance of 14 methods
from the test period. Evaluation metrics and computational time (in
minutes) for all methods are presented in Table \ref{tab:tab-results}.
They are ordered by Quantile Bias. The five models identified above have
a Quantile Bias of 0.014 or less, which is substantially lower than the
next group of forecasts with Quantile Biases of 0.037 and above, ETS
being the only exception with a value of 0.019.

One more thing to notice is that the ADAM-iETSX model with rounded up
quantiles did not perform better than the simpler one with continuous
ones (Table \ref{tab:tab-results}). This implies that the rounding is
not necessary in general, but if integer values are needed (for example,
to decide how many nurses to have), then using the continuous model and
then rounding up the quantiles could be considered a reasonable
strategy.

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/time-accuracy-1}

}

\caption{Running time vs. forecast performance}\label{fig:time-accuracy}
\end{figure}

Finally, Figure \ref{fig:time-accuracy} reports the forecast
performance of each approach versus the computational time required to
generate the forecast for a given forecasting horizon of 48 hours,
presented in Table \ref{tab:tab-results}. The X-axis shows the speed of
each method presented as slow, moderate or fast. We observe that there
is no clear association between speed and accuracy improvement of the
models used in this study.

We found that most of the methods considered in this paper fall into the
fast category with a different range of performance, depending on the
type of error measure. Among the fastest methods, ETS is very
competitive when assessed using Quantile Bias and RMSE. Moreover,
Poisson-1 and Poisson-2 provide accurate forecasts when evaluated by
RMSE and Pinball.

Figure \ref{fig:time-accuracy} shows that NOtr-2 is the fastest method
that provides consistent accuracy assessed via all three accuracy
measures. The figure also indicates that while the \(t\) family GAMLSS
model (Ttr-2) is the slowest method, it has a very good performance
across the three presented accuracy measures. Therefore, if the speed is
not a major constraint when generating the forecasts, one may employ
this approach to generate forecasts.

\textcolor{blue}{The benefits of the proposed model and its managerial aspects include
the use of probabilistic forecast to inform decision making and provide
insights on the running time of each model to generate forecasts. Probabilistic forecasts contain all potential future outcomes and help
planners to achieve a more efficient decision making by not only allocate a probability to the most likely outcome but also to unlikely
outcomes such as extremely high or low arrivals in the Emergency Departments. The running time could be also an important aspects for managers depending on the frequency of generating forecasts. If the frequency of forecast generation is high, one may employ models such as
Poisson regression (Poisson 1). If the frequency is lower and running time is not a big concern, a model like GAMLSS (Ttr-2) should be used.}

\textcolor{blue}{Probabilistic forecasts can be considered as a useful tool for planners to manage the risk of under and over resource allocation, which consequently may reduce both costs and risks of service delivery in Emergency Departments.}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Short-term forecasting of arrivals at emergency departments is an
important element of hospital staff and resource management.
Furthermore, due to the asymmetric impact of an excess resource
shortage, especially in emergency departments, quantifying forecast
uncertainty is also of value as it enables planners to manage associated
risks. In this paper, we have developed methods for producing a
probabilistic forecast of hourly arrivals up to 48 hours ahead,
comparing different state-of-the-art approaches.

Two approaches produced highly accurate, calibrated probabilistic
forecasts: a time series model and a model based on distributional
regression. The first is ADAM-iESTX, an extension of exponential
smoothing incorporating multiple seasonalities, explanatory variables
and assuming a Gamma predictive distribution. The second, labeled
NOtr-2, regressed the two parameters of a truncated (at zero) normal
distribution on the date and time features and temperature. Both
approaches produced calibrated probabilistic forecasts, but the point
prediction produced by ADAM-iESTX had a lower RMSE than NOtr-2, while
NOtr-2 produced forecasts with a lower Pinball score. This suggests that
the latter may be preferred if the whole distribution is used in
decision-making.

Having compared the performance of a wide range of methods, we make the
following observations: the choice of distribution assumed for
probabilistic forecasts and choice of model features are as if not more
important than the type of model employed; methods based on quantile
regression, which do not assume a parametric distribution for forecasts,
do not perform as well as those based on parametric distributions; and
the best performing models handled the non-negative and skewed nature of
the data automatically without the need for post-processing. These
observations reflect the characteristics of the data, which is
representative of ED arrivals, but determining the extent to which they
generalise is beyond the scope of this paper. Furthermore, methods based
on continuous-valued distributions are not adversely affected by the
fact that the data are integer-valued. Rounding up predictive quantiles
to the next integer does not make predictions worse.

Finally, we have found that out-of-the-box models, which require minimal
tuning or manual development, do not perform as well as well-considered
statistical approaches. The popular TBATS, Prophet and Gradient Boosting
Machine algorithms perform poorly compared to ADAM-iETSX and NOtr-2, and
even the benchmarks. Of the models requiring a modest amount of user
input and expertise, exponential smoothing (ETS) was found to perform
well. ETS produces reasonably well-calibrated forecasts, in contrast to
the poorly calibrated benchmarks and has highly accurate point
forecasts. However, its probabilistic forecasts were considerably worse
than NOtr-2 in terms of Pinball score.

\textcolor{blue}{
The dataset used in this study does not include the COVID-19 pandemic. During COVID-19, the dynamic of ED has changed. That means any forecasting model that is used during such periods to forecast ED arrivals needs to be modified to reflect those changes for that specific period. However, we expect that if the COVID period is taken into account by the models (or interpolated), the models should perform similarly.}

Probabilistic forecasting opens the door to more sophisticated resource
management in healthcare settings by providing decision-makers with
uncertainty information and enabling quantitative risk management.
Linking forecasts of arrivals with upstream (ambulance call-outs) and
downstream (length of stay, medical outcomes) analytics offers an
opportunity to improve forecasting skills and may also be necessary to
maximise benefits through more holistic decision-making.

Further research is required to investigate the practical benefits of
probabilistic forecasts in healthcare and how they can inform planning
and decision making. This may require employing discrete event simulation or
application to the newsvendor problems. While this study has focused on
hourly short-term forecasting, producing longer-term daily forecast
(e.g.~180-270 days ahead) is often required by planners to support
winter planning in ED and Ambulance services which requires more
investigation. Moreover, more research is needed in the forecasting of
other important variables such as length of stay, bed occupancy and
waiting time, in addition to patient arrivals and admissions. This may
require considering the dynamics among various services, including
General Practitioners, Emergency Departments, Ambulance and Fire \&
Rescue services.

\hypertarget{appendices}{%
\section{Appendices}\label{appendices}}

\hypertarget{quantilesceiling}{%
\subsection{Quantiles of rounded up random variables}\label{quantilesceiling}}

Before proceeding with the proof, we need to define the quantiles of the
continuous and rounded up random variables:
\begin{equation} \label{eq:quantCeil1}
    P \left(y_t < k \right) = 1 - \alpha ,
\end{equation} and \begin{equation} \label{eq:quantCeil2}
    P \left(\lceil y_t \rceil \leq n \right) \geq 1 - \alpha ,
\end{equation} where \(n\) is the quantile of the distribution of rounded
up values (the smallest integer number that satisfies the inequality
\eqref{eq:quantCeil2}) and \(k\) is the quantile of the continuous
distribution of the variable.

In order to prove that \(n = \lceil k \rceil\), we need to use the
following basic property: \begin{equation} \label{eq:quantCeil3}
    \lceil y_t \rceil \leq n \iff  y_t \leq n,
\end{equation} which means that the rounded up value will always be less
than or equal to \(n\) if and only if the original value is less than or
equal to \(n\). Taking into account \eqref{eq:quantCeil3}, the probability
\eqref{eq:quantCeil2} can be rewritten as:
\begin{equation} \label{eq:quantCeil4}
    P \left(y_t \leq n \right) \geq 1 - \alpha .
\end{equation}

Note also that the following is true:
\begin{equation} \label{eq:quantCeil5}
    P \left(\lceil y_t \rceil \leq n-1 \right) = P \left(y_t \leq n-1 \right) < 1 - \alpha .
\end{equation}

Taking the inequalities \eqref{eq:quantCeil1}, \eqref{eq:quantCeil2},
\eqref{eq:quantCeil4} and \eqref{eq:quantCeil5} into account, the
following can be summarised: \begin{equation} \label{eq:quantCeil6}
    P \left(y_t \leq n-1 \right) < P \left(y_t < k \right) \leq P \left(y_t \leq n \right) ,
\end{equation} which is possible only when \(k \in (n-1, n]\), which means
that \(\lceil k \rceil = n\). So the rounded up quantile of continuous
random variable \(y_t\) will always be equal to the quantile of the
descritised value of \(y_t\):
\begin{equation} \label{eq:ceilingAndQuantiles1}
    \left \lceil Q_\alpha(y_t) \right \rceil = Q_\alpha \left(\lceil y_t \rceil \right) .
\end{equation}

It is also worth noting that the same results can be obtained with the
floor function instead of ceiling, following the same logic. So the
following equation will hold for all \(y_t\) as well:
\begin{equation} \label{eq:floorAndQuantiles1}
    \left \lfloor Q_\alpha(y_t) \right \rfloor = Q_\alpha \left(\lfloor y_t \rfloor \right) .
\end{equation}

\renewcommand\refname{References}
\bibliography{mybibfile.bib}

\end{document}